---
title: "exploring"
author: "NAME"
date: "DATE"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readr)
LHK <- read_csv("C:/Users/Jim/Downloads/LHK.csv")
View(LHK)
```

Load the necessary libraries
```{r}
library(tidyverse)
```

Take a look inside your dataset 

```{r}
LHK
```


### Variation

Perform an analysis of the variation in the "neighbourhood" column. 

```{r}
ggplot(data = LHK) + 
  geom_bar(mapping = aes(x = neighbourhood)) + 
  coord_flip()
```

* Which values are the most common? Why?
The Yau Tsim Mong neighborhood has the most listings on AirBNB. This is because this neighbourhood is actually the biggest in Hong Kong. 

* Which values are rare? Why? Does that match your expectations?
There are a lot of neighbourhoods with relatively very few listings, however this is not unusual because the # of airBNB listings is mostly proportional to the population of the neighbourhood in Hong Kong. 

* Can you see any unusual patterns? What might explain them?
No. 


Perform an analysis of the variation in the "room_type" column. 

```{r}
ggplot(data = LHK) + geom_bar(mapping = aes(x = room_type, color = neighbourhood))
```

* Which values are the most common? Why?
The most common value here is the private room "room_type". This is because of Hong Kong's very urban setting. 

* Which values are rare? Why? Does that match your expectations?
The hotel room value is very rare, and this does match my expectations because I would assume that most hotels want to list their rooms seperately/ 

* Can you see any unusual patterns? What might explain them? 
I think that an unexpected trend was for the private room to have the most listings but I think this is due to the distinct environment of Hong Kong and how many skyscrapers there are with little land on the island. 



Perform an analysis of the variation in the "price" column. Make sure to explore different "binwidth" values in your analysis.

```{r} 
ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = price), binwidth = 150)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = price), binwidth = 100)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = price), binwidth = 200)

```

* Which values are the most common? Why?
A value around ~$400 a night seems to be the most common, however it is hard to visualize due to extreme outliers (I am pretty sure these outliers are not mistakes though, just very expensive AirBNB rentals)

* Which values are rare? Why? Does that match your expectations?
I would expect Hong Kong prices to be a little bit above average, but I bet that the price depends on the room_type variable as private rooms are bound to be less expensive than whole apartments or houses. 

* Can you see any unusual patterns? What might explain them? 
There are many outliers to the right, and one is ever visible upwards of 80k. This might be real or it could be fake, but there are also many others upwards of 10,000. I think this can be explained by the sheer price/sqft of properties in the middle of Hong Kong. For the properties listed as a whole house/apartment, if they have alot of square footage they will sell for a fortune on AirBNB. 



Perform an analysis of the variation in the "minimum_nights" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = minimum_nights), binwidth = 1)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = minimum_nights), binwidth = 5)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = minimum_nights), binwidth = 30)
```

* Which values are the most common? Why?
The value that is clearly the most common is simply one night. Every value between 1-7(AKA one week) is very commmon as well. This is because the sellers on AirBNB do not want to have to burden the customers with staying in a rental for too long, because this will dissuade many prospective buyers. 

* Which values are rare? Why? Does that match your expectations?
The rare values are basically any value over one month. This matches my expectations because people want to be able to rent their property to the largest audience possible and there are probably less people looking the longer the time interval is. This histogram shows what is called a "right-skewed" distribution in statistics. 


* Can you see any unusual patterns? What might explain them?
I do not see any unusual patterns in this histogram. I feel as if most of this was to be expected. 



Perform an analysis of the variation in the "number_of_reviews" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = number_of_reviews),binwidth = 1)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = number_of_reviews),binwidth = 10)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = number_of_reviews),binwidth = 50)
```

* Which values are the most common? Why?
It looks as if the most common number of reviews is between zero and one. This is because AirBNB's business appeals to first time renters who may not have experience renting out a home before. 

* Which values are rare? Why? Does that match your expectations?
Any number of reviews over about 75 seems pretty rare to me, however there are still a good amount of homes with this many reviews. This "right-skewed" distribution was to be expected due to the fact that AirBNB really appeals to the low-scale renter, while also providing a platform for someone to rent out their property hundreds of times and receive a lot of reviews. 


* Can you see any unusual patterns? What might explain them?
Nothing unusual. 


Perform an analysis of the variation in the "availability_365" column. Make sure to explore different "binwidth" values in your analysis.

```{r}
ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = availability_365), binwidth = 1)

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = availability_365), binwidth = 5) 

ggplot(data = LHK) + 
  geom_histogram(mapping = aes(x = availability_365), binwidth = 10)
```

* Which values are the most common? Why?
The most common value is 365, because this signals that this rental property is available for the whole year, which is pretty common because Hong Kong is not too seasonal of a destination. 

* Which values are rare? Why? Does that match your expectations?
There are not really any "rare" values here so to speak, because I feel that the values 1-364 all seem roughly equally likely (with the exception of the two humps)

* Can you see any unusual patterns? What might explain them?
I think it is unusual to see a quadmodal (4 humps) histogram distribution, but I think that it can be explained using the seasonal context of AirBNB. 

Questions (Part 2) 
Q1: What seems to be the most common name (of a person) in the city you selected?

```{r}
count(LHK, host_name) %>% 
  arrange(desc(n))
```
Answer: The most common name of a person on AirBNB hosting in Hong Kong is Jovee. 

Q2: Do the number of reviews affect the price of the AirBNB? How? Why do you think this is? 

```{r}
LHK_filter <- filter(LHK, price < 2000)
ggplot(data = LHK_filter) +
  geom_point(mapping = aes(x = number_of_reviews, y = price)) + 
  geom_smooth(mapping = aes(x = number_of_reviews, y = price))  
  
```

Answer: There is really no effect of number of reviews on price of AirBNBs in Hong Kong. I think this is because there are many other factors in Hong Kong AirBNBs that matter a lot more. 

Q3: What type of room tends to have the highest AirBNB price? 

```{r}
ggplot(data = LHK_filter) + 
  geom_histogram(mapping = aes(x = price, fill = room_type), binwidth = 50)
```
Answer: the entire home/apartment room type tends to have the highest AirBNB price. 

Q4: What neighborhoods tend to have the highest AirBNB price 

```{r}
ggplot(data = LHK_filter) + 
  geom_boxplot(mapping = aes(x = neighbourhood, y = price)) + 
  coord_flip()

```
Answer: The Kwai Tsing neighborhood of Hong Kong has the highest median AirBNB price

Q5: Suppose you could purchase a property in the city you selected, and that you could rent it to others as an AirBNB. In what neighborhood would you purchase your property? Why? 

```{r}
#want to calculate the difference between the property prices and median AirBNB price in the neighborhood; Do some research 


```


Answer: I would purchase a home in the Southern neighborhood because from the period 2010-2019, home prices were steadily increasing and outpacing the rest of Hong Kong, which can be expected to continue.  


Part 3 
```{r}
LHK_by_neighborhood <- LHK %>% 
  group_by(neighbourhood) 

LHK_by_neighborhood
```

Use your dataset to find what the average price/night is in the neighborhood you selected
```{r}
ggplot(data = LHK_filter) + 
  geom_boxplot(mapping = aes(x = neighbourhood, y = price)) + 
  coord_flip()
```
Answer: The average price per night in my neighbourhood(Southern) is about 750 dollars per night. 

Use your dataset to find what the average number of available nights per year is in the neighborhood you selected. 
```{r} 
ggplot(data = LHK_filter) + 
  geom_boxplot(mapping = aes(x = neighbourhood, y = availability_365)) + 
  coord_flip()
```
Answer: the average number of available nights in the Southern neighborhood is about 260 nights. 

### Calculations of break even price 

It will take approximately 7 years to break even on AirBNB using the average number of nights available and ~700 dollars per night (my property is a little bit below average in the neighborhood at only 1.2 million dollars). In conclusion, Hong Kong's Southern neighborhood is not the best place to try to build an AirBNB business. This is because Hong Kong is already dominated by the condo market that buying homes is so expensive. 

```{r}
7*700*260
```

(ALL DOLLAR VALUES ARE USD)





